CF Registry Analysis

### 2024-04-29 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
survival code cleanup

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
source("/Users/timvigers/GitHub/shared-resources/Visualization/forest_plots.R")
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# All data
encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual <- annual %>% arrange(eDWID, ReviewYear)
# Add age
ages <- annual %>%
  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID, reviewyear) %>%
  summarise(
    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
      decreasing = TRUE
    )[1])),
    Age = Age_YrEnd[1], .groups = "drop"
  )
encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
encounter$`Age Group` <- cut(encounter$Age,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>%
  filter(cfrd_status == 3) %>%
  group_by(eDWID) %>%
  filter(row_number() == 1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_cffpr,
  by = c("eDWID", "Age in 2018")
)
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
# Ordered factors
encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
  levels = c("No", "Yes"), ordered = T
)
# Remove those missing age
encounter <- encounter %>% filter(!is.na(Age))
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc <- function(df = encounter, year, py = 1000) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, is.na(first_cfrd_cffpr) |
      first_cfrd_cffpr >= year) %>%
    group_by(eDWID) %>%
    filter(row_number() == 1) %>%
    select(eDWID, `Age Group`, cfrd_cffpr)
  inc <- inc %>%
    group_by(cfrd_cffpr) %>%
    count(`Age Group`, .drop = F)
  inc <- inc[!is.na(inc$`Age Group`), ]
  # Calculations
  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
  total_at_risk <- sum(inc$n)
  total_inc <- new_cases / (total_at_risk / py)

  new_cases_10 <-
    sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 <- new_cases_10 / (at_risk_10 / py)

  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
    inc$`Age Group` == "10 - 18"])
  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
    inc$`Age Group` == "19 - 30"])
  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
    inc$`Age Group` == "> 30"])
  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 <- new_cases_30 / (at_risk_30 / py)
  # Nice results
  list(
    `Year` = year,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age > 30` = new_cases_30,
    `Total At Risk Age > 30` = at_risk_30,
    `Total Incidence Age > 30` = inc_30
  )
}
incidence <- lapply(2003:2018, function(y) {
  year_inc(year = y)
})
incidence <- data.frame(do.call(rbind, incidence), check.names = F)
incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
kable(incidence, digits = 3)
```

In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(
    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
    .groups = "drop"
  ) %>%
  kable(., digits = 3)
```

```{r}
#| label: tbl-prevalence_cffpr_year
#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear, `Age Group`) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear, `Age Group`) %>%
  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
  kable(., digits = 3)
```

In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-screening_ogtt
#| tbl-cap: "Yearly OGTT Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year |
      is.na(first_cfrd_cffpr)) %>%
    mutate(ogtt = !is.na(ogttfast_bloodglucose) |
      !is.na(twohour_bloodglucose)) %>%
    group_by(eDWID) %>%
    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(ogtt) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$ogtt == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

```{r}
#| label: tbl-screening_a1c
#| tbl-cap: "Yearly HbA1c Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year |
      is.na(first_cfrd_cffpr)) %>%
    mutate(a1c = !is.na(hgba1c)) %>%
    group_by(eDWID) %>%
    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(a1c) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$a1c == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.

@@

@@

@@

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual", 
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
# Labels
label(s) <- labels[colnames(s)]
```

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation",  "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

@@

```{r}
s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
  data = s, cut = c(2008, 2013, 2018), episode = "tgroup"
)
s_split$tgroup <- factor(s_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_int <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*tgroup")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
  cox_res_split_int <- coxph(new_f_int, data = s_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression() %>%
    add_n(location = "level") %>%
    add_nevent(location = "level") %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split_int, new_f_emm,
    nuisance = xvars[-which(xvars == v)]
  )
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_model(cox_res_split,
    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
    include = xvars[-which(xvars == v)]
  ) %>%
    print()
  cat("\n")
  cat("\n")
  # Print EMMs
  print(kable(emm, digits = 3))
  cat("\n")
  cat("\n")
  # Plot marginal means
  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
    print()
  cat("\n")
  cat("\n")
}))
```

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- ### Vx770 -->

<!-- ```{r} -->
<!-- f <- update(f, . ~ . + A_Vx770) -->
<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
<!-- # Plot survival -->
<!-- ggsurvplot(survfit(cox_res), -->
<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
<!--   xlim = c(2003, 2018), break.x.by = 1 -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- # Diagnostics -->
<!-- # test.ph= cox.zph(cox_res) -->
<!-- # ggcoxzph(test.ph) -->
<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
<!-- ``` -->

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Mycobacterium -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- ### Vx809 -->

<!-- ```{r} -->
<!-- f <- update(f, . ~ . + A_VX809comb) -->
<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
<!-- # Plot survival -->
<!-- ggsurvplot(survfit(cox_res), -->
<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
<!--   xlim = c(2003, 2018), break.x.by = 1 -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- # Diagnostics -->
<!-- # test.ph= cox.zph(cox_res) -->
<!-- # ggcoxzph(test.ph) -->
<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
<!-- ``` -->

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Mycobacterium -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- t -->

>>>>

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
library(rprev)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Find each person's entry date
annual = annual %>% group_by(eDWID) %>%
  mutate(entry_date = min(ReviewYear))

@@

# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
# years following are CFRD+
annual$CFRD <- NA
annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
# For each person, filter out all the years after first CFRD diagnosis
annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
# Remove those under 10
annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
annual$age_group = droplevels(annual$age_group)
# Add formatted columns for survival analysis
annual = annual %>% 
  mutate(tstop = ReviewYear) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year",
  tstart = "Interval start", tstop = "Interval stop",
  age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"

@@

```{r}
t2004 = annual %>% select()
```

```{r}
# Use flexsurv to fit a survival model that can be passed to rprev
llog <- flexsurv::flexsurvreg(Surv(tstop, as.numeric(CFRD)==2) ~ 1, 
                              data=prevdata, dist='llogis')
prev_llog <- prevalence(index='2004-01-01', 
                        num_years_to_estimate=1, 
                        data=annual, 
                        inc_formula = entrydate ~ 1,
                        surv_model=llog,
                        death_column = "first_cfrd",
                        incident_column = "first_cfrd",
                        N_boot = 100)
```

@@

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ 1,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ age_group,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, as.numeric(CFRD)==2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, id = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

@@

```{r}
annual_split <- survSplit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_int <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*tgroup")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  cox_res_split_int <- coxph(new_f_int, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression() %>%
    add_n(location = "level") %>%
    add_nevent(location = "level") %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split_int, new_f_emm,
    nuisance = xvars[-which(xvars == v)]
  )
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_model(cox_res_split,
    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
    include = xvars[-which(xvars == v)]
  ) %>%
    print()
  cat("\n")
  cat("\n")
  # Print EMMs
  print(kable(emm, digits = 3))
  cat("\n")
  cat("\n")
  # Plot marginal means
  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
    print()
  cat("\n")
  cat("\n")
}))
```
